---
title: "Pre-processing script - Merge all participants files and aggregate measures"
author: "Zoé Bollen"
date: "`r Sys.Date()`"
output: html_document
---

```{r package, include=FALSE, echo=FALSE}
# install package librarian if needed
if (!("librarian" %in% rownames(installed.packages()))) {
  install.packages("librarian")
}

# install / load required packages
librarian::shelf(
  dplyr,
  readr
)
```

# Merge all files

## Load and merge gaze-behavioural data of all participants

```{r warning=FALSE}
#set the path
PreprocessMainDir <- '../preprocessed data'
AllSubDirFix <- "Each_Part_Fix_Outputs"

setwd(file.path(PreprocessMainDir, AllSubDirFix))

#load and merge all participants data
files_to_upload <- list.files(pattern="*.csv")
theData_list <-lapply(files_to_upload, read.csv)
All_Part_Data<-bind_rows(theData_list)
```

## Data check

Check if all participants data are available by printing the number of participants and their individual participant codes.

```{r}
length(unique(All_Part_Data$part_ID))
unique(All_Part_Data$part_ID)
```

## Save merged data into a single file

```{r warning=FALSE}
setwd(file.path(PreprocessMainDir))
write.csv(All_Part_Data, file = paste("All_Part_FixBehav_Data.csv",sep=""), row.names = FALSE)
```

# Aggregate measures

```{r warning=FALSE, message=FALSE}
#load file
setwd(file.path(PreprocessMainDir))

All_Part_Data <- read_csv(file = "All_Part_FixBehav_Data.csv")
```

## Unhealthy food craving

```{r}
#ADD A COLUMN WHICH INDICATES THE MEAN CRAVING FOR UNHEALTHY FOOD
All_Part_Data$unhealthy_craving <- rowMeans(All_Part_Data[, c("sugar_craving", "fat_craving")])
```

## Number of fixations

```{r message=FALSE}
# CALCULATE NUMBER OF FIXATIONS PER AOI
NbFix <- All_Part_Data %>%
  group_by (
    part_ID,
    sugar_craving,
    fat_craving,
    unhealthy_craving,
    trial_id,
    Left_Picture,
    Right_Picture,
    accuracy,
    response_time,
    Congruence,
    AOI_visited) %>%
  summarise(n = n()) %>%
  ungroup() 

# TRANSPOSE CASES TO COLUMNS TO HAVE ONE LINE PER TRIAL AND ONE COLUMN PER NUMBER OF FIXATIONS ON EACH AOI
fixation_count <- tidyr::pivot_wider(NbFix, names_from = AOI_visited, values_from = n)

# RENAME THE COLUMNS
names(fixation_count)[names(fixation_count) == "Healthy"] <- "FixCount_Healthy"
names(fixation_count)[names(fixation_count) == "Sugary"] <- "FixCount_Sugary"
names(fixation_count)[names(fixation_count) == "Salty"] <- "FixCount_Salty"

#NOTE AN ABSENCE OF FIXATION ON ONE AOI (NA) BY 0 FIXATION FOR RELEVANT TRIALS (TRIAL 1 TO 40 FOR SALTY AOI, TRIAL 41 TO 80 FOR SUGARY AOI)
fixation_count$FixCount_Healthy[is.na(fixation_count$FixCount_Healthy)] <- 0 
fixation_count$FixCount_Salty[is.na(fixation_count$FixCount_Salty) & fixation_count$trial_id %in% 1:40] <- 0 
fixation_count$FixCount_Sugary[is.na(fixation_count$FixCount_Sugary) & fixation_count$trial_id %in% 41:80] <- 0 

#CALCULATE AND ADD ONE COLUMN INDICATING THE TOTAL NUMBER OF FIXATIONS PER TRIAL
fixation_count$FixCount_Total <- round(rowSums(fixation_count[, c("FixCount_Healthy", "FixCount_Salty", "FixCount_Sugary")], na.rm = TRUE), 2)

#CALCULATE AND ADD COLUMN INDICATING THE RATIO OF FIXATIONS FOR UNHEALTHY FOOD
fixation_count$Rel_FixCount_Salty <- round(fixation_count$FixCount_Salty/fixation_count$FixCount_Total, 2)
fixation_count$Rel_FixCount_Sugary <- round(fixation_count$FixCount_Sugary/fixation_count$FixCount_Total, 2)
fixation_count$Rel_FixCount_Unhealthy <- round(
  ifelse(!is.na(fixation_count$FixCount_Sugary),
    fixation_count$FixCount_Sugary / fixation_count$FixCount_Total,
    fixation_count$FixCount_Salty / fixation_count$FixCount_Total),
    2
  )
```

## Dwell time

```{r message=FALSE}
# CALCULATE DWELL TIME PER AOI (while keeping information concerning part and trial)
DwellTime <- All_Part_Data %>%
  group_by (
    part_ID,
    sugar_craving,
    fat_craving,
    unhealthy_craving,
    trial_id,
    Left_Picture,
    Right_Picture,
    accuracy,
    response_time,
    Congruence, 
    AOI_visited)  %>%
  summarise(DT = sum(duration))  %>%
  ungroup()

# TRANSPOSE CASES TO COLUMNS TO HAVE ONE LINE PER TRIAL AND ONE COLUMN PER DWELL TIME ON EACH AOI
dwell_time <- tidyr::pivot_wider(DwellTime, names_from = AOI_visited, values_from = DT)

# RENAME THE COLUMNS
names(dwell_time)[names(dwell_time) == "Healthy"] <- "DT_Healthy"
names(dwell_time)[names(dwell_time) == "Salty"] <- "DT_Salty"
names(dwell_time)[names(dwell_time) == "Sugary"] <- "DT_Sugary"

#NOTE AN ABSENCE OF FIXATION ON ONE AOI (NA) BY 0 MS FOR RELEVANT TRIALS (TRIAL 1 TO 40 FOR SALTY AOI, TRIAL 41 TO 80 FOR SUGARY AOI)
dwell_time$DT_Healthy[is.na(dwell_time$DT_Healthy)] <- 0 
dwell_time$DT_Salty[is.na(dwell_time$DT_Salty) & dwell_time$trial_id %in% 1:40] <- 0 
dwell_time$DT_Sugary[is.na(dwell_time$DT_Sugary) & dwell_time$trial_id %in% 41:80] <- 0 

#CALCULATE AND ADD ONE COLUMN INDICATING THE TOTAL DWELL TIME PER TRIAL
dwell_time$DT_Total <- round(rowSums(dwell_time[, c("DT_Healthy", "DT_Salty", "DT_Sugary")], na.rm = TRUE), 2)

#CALCULATE AND ADD COLUMN INDICATING THE RELATIVE DWELL TIME FOR UNHEALTHY FOOD
dwell_time$Rel_DT_Salty <- round(dwell_time$DT_Salty/dwell_time$DT_Total, 2)
dwell_time$Rel_DT_Sugary <- round(dwell_time$DT_Sugary/dwell_time$DT_Total, 2)
dwell_time$Rel_DT_Unhealthy <- round(
  ifelse(!is.na(dwell_time$DT_Sugary), 
         dwell_time$DT_Sugary / dwell_time$DT_Total, 
         dwell_time$DT_Salty / dwell_time$DT_Total),
  2
)
```

## First fixation

```{r message=FALSE}
# CREATE A VARIABLE INDICATING THE ORDER OF FIXATIONS WITHIN EACH TRIAL
All_Part_Data_with_fixation_order <- All_Part_Data %>%
  arrange(part_ID, trial_id, timestamp) %>% 
  group_by(part_ID, trial_id) %>%           
  mutate(fixation_order = row_number()) %>% 
  ungroup()     

# CALCULATE MIN FIXATION ORDER PER AOI
FirstFix <- All_Part_Data_with_fixation_order %>%
  group_by (
    part_ID,
    sugar_craving,
    fat_craving,
    unhealthy_craving,
    trial_id,
    Left_Picture,
    Right_Picture,
    accuracy,
    response_time,
    Congruence,
    AOI_visited)  %>%
  summarise(Min_fix_order = min(fixation_order)) %>%
  ungroup()

# DETERMINE WHETHER THE FIRST FIXATION WAS MADE ON THE AOI
FirstFix <- FirstFix %>%
  mutate(
    First_Fix = case_when(
      Min_fix_order == 1 ~ 1,  
      Min_fix_order > 1 ~ 0,                                   
      TRUE ~ NA_real_)
  ) %>%
  select(-Min_fix_order)

#PIVOT DATASET
first_fix <- tidyr::pivot_wider(FirstFix, names_from = AOI_visited, values_from = First_Fix)

# RENAME THE COLUMNS
names(first_fix)[names(first_fix) == "Healthy"] <- "first_fix_Healthy"
names(first_fix)[names(first_fix) == "Salty"] <- "first_fix_Salty"
names(first_fix)[names(first_fix) == "Sugary"] <- "first_fix_Sugary"

#DETERMINE WHETHER THE FIRST FIXATION WAS MADE ON UNHEALTHY FOOD
first_fix$first_fix_Unhealthy <- 
  ifelse(!is.na(first_fix$first_fix_Sugary), 
         first_fix$first_fix_Sugary, 
         first_fix$first_fix_Salty)

#KEEP ONLY RELEVANT VARIABLES
first_fix <- select(
  first_fix,
  part_ID,
  sugar_craving,
  fat_craving,
  unhealthy_craving,
  trial_id,
  first_fix_Salty,
  first_fix_Sugary, 
  first_fix_Unhealthy
)
```

# Merge measures in one single dataset

```{r}
#longer format
list_measures_AOI <- list(DwellTime, FirstFix, NbFix)
All_Measures_PerAOI <- Reduce(function(x, y) merge(x, y, all=TRUE), list_measures_AOI) 

#wider format
list_measures_Trial <- list(dwell_time, first_fix, fixation_count)
All_Measures_PerTrial <- Reduce(function(x, y) merge(x, y, all=TRUE), list_measures_Trial)  
```

# Data reduction procedure

## Exclude trials with uncorrect responses

Remove all trials in which participants did not answer correctly to the probe

```{r}
# Count the number of trials where accuracy = 0
num_uncorrect_trials <- All_Measures_PerTrial %>%
  group_by(trial_id) %>%
  summarise(uncorrect = any(accuracy == 0)) %>%
  filter(uncorrect) %>%
  nrow()

# Print the result
print(num_uncorrect_trials)

# Keep only rows where responses are correct
All_Measures_PerTrial <- subset(All_Measures_PerTrial, accuracy == 1)
All_Measures_PerAOI <- subset(All_Measures_PerAOI, accuracy == 1)
```

## Exclude trials when response time \< 200ms

```{r}
# Filter to find trials with decision < 200ms
trials_RT_lower_200ms <- All_Measures_PerTrial %>%
  filter(response_time < 200)

# Print the result
print(trials_RT_lower_200ms)

# Exclude those trials
All_Measures_PerTrial <- All_Measures_PerTrial %>%
  filter(response_time >= 200)
All_Measures_PerAOI <- All_Measures_PerAOI %>%
  filter(response_time >= 200)
```

## Exclude trials when response time \> 2000ms

```{r}
# Filter to find trials with decision > 2000ms
trials_RT_higher_2000ms <- All_Measures_PerTrial %>%
  filter(response_time > 2000)

# Print the result
print(trials_RT_higher_2000ms)

# Exclude those trials
All_Measures_PerTrial <- All_Measures_PerTrial %>%
  filter(response_time <= 2000)
All_Measures_PerAOI <- All_Measures_PerAOI %>%
  filter(response_time <= 2000)
```

## Exclude participants with half of trials without fixations

```{r}
# Create a summary of the number of trials per subject
subject_trial_count <- All_Measures_PerTrial %>%
  group_by(part_ID) %>%
  summarize(trial_count = n()) %>%
  ungroup()

# Filter to find subjects with fewer than 40 trials
subjects_with_fewer_than_40_trials <- subject_trial_count %>%
  filter(trial_count < 40)

# View the subjects with fewer than 40 trials
print(subjects_with_fewer_than_40_trials)

# Filter to find subjects with at least half of the trials
subjects_with_40_or_more_trials <- subject_trial_count %>%
  filter(trial_count >= 40)

# Exclude the participant(s) 
All_Measures_PerTrial <- All_Measures_PerTrial %>%
  filter(part_ID %in% subjects_with_40_or_more_trials$part_ID)
All_Measures_PerAOI <- All_Measures_PerAOI %>%
  filter(part_ID %in% subjects_with_40_or_more_trials$part_ID)
```

# Calculate reaction time AB scores

## Calculate RT AB score for unhealthy versus healthy food

```{r message=FALSE}
# CALCULATE MEAN REACTION TIME PER CONGRUENCE
RT_PerCongruence <- All_Measures_PerTrial %>%
  group_by(part_ID, Congruence) %>%
  summarize(RT = mean(response_time)) %>%
  ungroup()

# RENAME THE CONGRUENCE VALUES
RT_PerCongruence <- RT_PerCongruence %>%
  mutate(Congruence = recode(Congruence, 
                              "0" = "Uncongruent",
                              "1" = "Congruent"))

# TRANSPOSE CASES TO COLUMNS TO HAVE ONE LINE PER TRIAL AND ONE COLUMN PER DWELL TIME ON EACH AOI
Reaction_Time <- tidyr::pivot_wider(RT_PerCongruence, names_from = Congruence, values_from = RT)

#CALCULATE THE REACTION TIME AB SCORE (POSITIVE VALUES INDICATED A BIAS TOWARDS UNHEALTHY FOOD AND NEGATIVE VALUES INDICATED A BIAS TOWARDS HEALTHY FOOD)
Reaction_Time$RT_AB_score <- round(Reaction_Time$Uncongruent - Reaction_Time$Congruent, 2)

#KEEP ONLY RELEVANT VARIABLES
Reaction_Time <- select(
  Reaction_Time,
  part_ID,
  RT_AB_score
)
```

## Calculate RT AB score for salty versus healthy food

```{r message=FALSE}
# CALCULATE MEAN REACTION TIME PER CONGRUENCE ONLY IN TRIALS WITH SALTY FOOD
RT_Salty_PerCongruence <- All_Measures_PerTrial %>%
  filter(trial_id >= 1 & trial_id <= 40) %>%
  group_by(part_ID, Congruence) %>%
  summarize(RT = mean(response_time)) %>%
  ungroup()

# RENAME THE CONGRUENCE VALUES
RT_Salty_PerCongruence <- RT_Salty_PerCongruence %>%
  mutate(Congruence = recode(Congruence, 
                              "0" = "Uncongruent",
                              "1" = "Congruent"))

# TRANSPOSE CASES TO COLUMNS TO HAVE ONE LINE PER TRIAL AND ONE COLUMN PER DWELL TIME ON EACH AOI
Reaction_Time_Salty <- tidyr::pivot_wider(RT_Salty_PerCongruence, names_from = Congruence, values_from = RT)

#CALCULATE THE REACTION TIME AB SCORE (POSITIVE VALUES INDICATED A BIAS TOWARDS UNHEALTHY FOOD AND NEGATIVE VALUES INDICATED A BIAS TOWARDS HEALTHY FOOD)
Reaction_Time_Salty$RT_Salty_AB_score <- round(Reaction_Time_Salty$Uncongruent - Reaction_Time_Salty$Congruent, 2)

#KEEP ONLY RELEVANT VARIABLES
Reaction_Time_Salty <- select(
  Reaction_Time_Salty,
  part_ID,
  RT_Salty_AB_score
)
```

## Calculate RT AB score for sugary versus healthy food

```{r message=FALSE}
# CALCULATE MEAN REACTION TIME PER CONGRUENCE ONLY IN TRIALS WITH SALTY FOOD
RT_Sugary_PerCongruence <- All_Measures_PerTrial %>%
  filter(trial_id >= 41 & trial_id <= 80) %>%
  group_by(part_ID, Congruence) %>%
  summarize(RT = mean(response_time)) %>%
  ungroup()

# RENAME THE CONGRUENCE VALUES
RT_Sugary_PerCongruence <- RT_Sugary_PerCongruence %>%
  mutate(Congruence = recode(Congruence, 
                              "0" = "Uncongruent",
                              "1" = "Congruent"))

# TRANSPOSE CASES TO COLUMNS TO HAVE ONE LINE PER TRIAL AND ONE COLUMN PER DWELL TIME ON EACH AOI
Reaction_Time_Sugary <- tidyr::pivot_wider(RT_Sugary_PerCongruence, names_from = Congruence, values_from = RT)

#CALCULATE THE REACTION TIME AB SCORE (POSITIVE VALUES INDICATED A BIAS TOWARDS UNHEALTHY FOOD AND NEGATIVE VALUES INDICATED A BIAS TOWARDS HEALTHY FOOD)
Reaction_Time_Sugary$RT_Sugary_AB_score <- round(Reaction_Time_Sugary$Uncongruent - Reaction_Time_Sugary$Congruent, 2)

#KEEP ONLY RELEVANT VARIABLES
Reaction_Time_Sugary <- select(
  Reaction_Time_Sugary,
  part_ID,
  RT_Sugary_AB_score
)
```

# Create and save datasets for analyses

## Dataset containing one row per participant

```{r message=FALSE}
#Create a new dataset containing one row per participant and calculating the mean values for each measures
All_Measures_PerPart <- All_Measures_PerTrial %>%
  group_by(part_ID, sugar_craving, fat_craving, unhealthy_craving) %>%
  summarize(
    Rel_DT_Sugary_mean = mean(Rel_DT_Sugary, na.rm = TRUE),
    Rel_DT_Sugary_sd = sd(Rel_DT_Sugary, na.rm = TRUE),
    Rel_DT_Salty_mean = mean(Rel_DT_Salty, na.rm = TRUE),
    Rel_DT_Salty_sd = sd(Rel_DT_Salty, na.rm = TRUE),
    Rel_DT_Unhealthy_mean = mean(Rel_DT_Unhealthy, na.rm = TRUE),
    Rel_DT_Unhealthy_sd = sd(Rel_DT_Unhealthy, na.rm = TRUE),
    FixCount_Total_mean = mean(FixCount_Total, na.rm = TRUE),
    FixCount_Total_sd = sd(FixCount_Total, na.rm = TRUE),
    Rel_FixCount_Sugary_mean = mean(Rel_FixCount_Sugary, na.rm = TRUE),
    Rel_FixCount_Sugary_sd = sd(Rel_FixCount_Sugary, na.rm = TRUE),
    Rel_FixCount_Salty_mean = mean(Rel_FixCount_Salty, na.rm = TRUE),
    Rel_FixCount_Salty_sd = sd(Rel_FixCount_Salty, na.rm = TRUE),
    Rel_FixCount_Unhealthy_mean = mean(Rel_FixCount_Unhealthy, na.rm = TRUE),
    Rel_FixCount_Unhealthy_sd = sd(Rel_FixCount_Unhealthy, na.rm = TRUE),
    first_fix_Sugary_mean = mean(first_fix_Sugary, na.rm = TRUE),
    first_fix_Sugary_sd = sd(first_fix_Sugary, na.rm = TRUE),
    first_fix_Salty_mean = mean(first_fix_Salty, na.rm = TRUE),
    first_fix_Salty_sd = sd(first_fix_Salty, na.rm = TRUE),
    first_fix_Unhealthy_mean = mean(first_fix_Unhealthy, na.rm = TRUE),
    first_fix_Unhealthy_sd = sd(first_fix_Unhealthy, na.rm = TRUE),
  ) %>%
  ungroup()

#Merge the reaction time AB score with the All_Measures_PerPart dataset
All_Measures_PerPart <- All_Measures_PerPart %>%
  left_join(Reaction_Time, by = "part_ID") %>%
  left_join(Reaction_Time_Salty, by = "part_ID") %>%
  left_join(Reaction_Time_Sugary, by = "part_ID")
```

## Save file containing all measures for analyses

```{r warning=FALSE}
setwd(file.path(PreprocessMainDir))

#Save one file containing one row per AOI visited
write.csv(All_Measures_PerAOI, file = paste("AllMeasures_ForAnalysis_PerAOI.csv",sep=""), row.names = FALSE)

#Save one file containing one row per trial
write.csv(All_Measures_PerTrial, file = paste("AllMeasures_ForAnalysis_PerTrial.csv",sep=""), row.names = FALSE)

#Save one file containing one row per part
write.csv(All_Measures_PerPart, file = paste("AllMeasures_ForAnalysis_PerPart.csv",sep=""), row.names = FALSE)
```
