---
title: "Analyses script - Descriptive statistics and regression analyses"
author: "Zoé Bollen"
date: "`r Sys.Date()`"
output: html_document
---

```{r package, include=FALSE, echo=FALSE}
# install package librarian if needed
if (!("librarian" %in% rownames(installed.packages()))) {
  install.packages("librarian")
}

# install / load required packages
librarian::shelf(
  dplyr,
  readr, 
  tidyr,
  ggplot2,
  lme4,
  sjPlot
)
```

## Load all measures files

```{r message=FALSE, warning=FALSE}
#set the path
PreprocessMainDir <- '../preprocessed data'
setwd(file.path(PreprocessMainDir))

#load file containing mean measures per AOI visited 
All_Measures_PerAOI <- read_csv(file = "AllMeasures_ForAnalysis_PerAOI.csv")

#load file containing measures per trial 
All_Measures_PerTrial <- read_csv(file = "AllMeasures_ForAnalysis_PerTrial.csv")

#load file containing mean measures per participant 
All_Measures_PerPart <- read_csv(file = "AllMeasures_ForAnalysis_PerPart.csv")
```

# Descriptive statistics

## Sample size

```{r}
#Number of participants after data reduction procedure
count_part <- count(All_Measures_PerPart)
print(count_part)
```

## Subjective craving

```{r warning=FALSE}
mean_sugary_craving <- mean(All_Measures_PerPart$sugar_craving)
mean_salty_craving <- mean(All_Measures_PerPart$fat_craving)

# histogram for sugary craving
ggplot(All_Measures_PerPart, aes(x=sugar_craving)) + 
  geom_histogram(color="black", fill="darkblue", binwidth = 1, alpha = 0.5) +
  labs(title = "Distribution of sugary craving",
       x = "Sugary craving",
       y = "Participant count") +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, 5, by = 1)) +
  scale_x_continuous(breaks = seq(0, 100, by = 10), limits = c(0, 100)) + 
  geom_vline(aes(xintercept = mean_sugary_craving), color = "grey", linetype = "dashed", size = 1) + 
  geom_text(aes(x = mean_sugary_craving, y = 5, label = paste("Mean =", round(mean_sugary_craving, 2))), 
            color = "darkgrey", vjust = 1, hjust = -0.25) + 
  theme(
    plot.title = element_text(hjust = 0.5) 
  )

# histogram for salty craving
ggplot(All_Measures_PerPart, aes(x=fat_craving)) + 
  geom_histogram(color="black", fill="darkblue", binwidth = 1, alpha = 0.5) +
  labs(title = "Distribution of salty craving",
       x = "Salty craving",
       y = "Participant count") +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, 5, by = 1)) +
  scale_x_continuous(breaks = seq(0, 100, by = 10), limits = c(0, 100)) + 
  geom_vline(aes(xintercept = mean_salty_craving), color = "grey", linetype = "dashed", size = 1) + 
  geom_text(aes(x = mean_salty_craving, y = 5, label = paste("Mean =", round(mean_salty_craving, 2))), 
            color = "darkgrey", vjust = 1, hjust = -0.25) + 
  theme(
    plot.title = element_text(hjust = 0.5) 
  )
```

## Reaction time

```{r}
All_Measures_PerTrial$Congruence<-as.character(All_Measures_PerTrial$Congruence)

# Boxplot for reaction time for congruent vs uncongruent trials
ggplot(All_Measures_PerTrial, aes(x = Congruence, y = response_time, fill = Congruence)) +
  geom_boxplot() +
  labs(x = "Congruence", y = "Response time") +
  theme_minimal() +
  theme(legend.position = "none") +
  ylim(0, 2000)
```

### Reaction time and craving

```{r message=FALSE}
# Scatter plot for Reaction time on unhealthy food and craving
ggplot(All_Measures_PerPart, aes(x = unhealthy_craving, y = RT_AB_score)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue") +
  labs(x = "Craving for unhealthy food",
       y = "Reaction time AB score for unhealthy food") +
  theme_minimal() +
  ylim(-200, 200) +
  xlim(0, 100)

```

## First fixation

### first fixation on healthy versus unhealthy food

```{r message=FALSE}
# Filter dataset to include only trials 1 to 40
Salty_trials <- All_Measures_PerAOI %>%
  filter(trial_id >= 1 & trial_id <= 40)

#Create the stacked bar plot
ggplot(Salty_trials, aes(x = factor(AOI_visited), fill = factor(First_Fix))) +
  geom_bar(position = "fill") +
  labs(x = "Area of Interest (AOI)", y = "Proportion", fill = "First Fixation") +
  scale_fill_manual(values = c("lightblue", "darkblue")) +
  theme_minimal()

# Filter dataset to include only trials 1 to 40
Sugary_trials <- All_Measures_PerAOI %>%
  filter(trial_id >= 41 & trial_id <= 80)

#Create the stacked bar plot
ggplot(Sugary_trials, aes(x = factor(AOI_visited), fill = factor(First_Fix))) +
  geom_bar(position = "fill") +
  labs(x = "Area of Interest (AOI)", y = "Proportion", fill = "First Fixation") +
  scale_fill_manual(values = c("lightblue", "darkblue")) +
  theme_minimal()
```

### first fixation on unhealthy food and craving

```{r message=FALSE}
# Scatter plot for First fixation on unhealthy food and craving
ggplot(All_Measures_PerPart, aes(x = unhealthy_craving, y = first_fix_Unhealthy_mean)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue") +
  labs(x = "Craving for unhealthy food",
       y = "First fixation on unhealthy food") +
  theme_minimal() +
  ylim(0, 1) +
  xlim(0, 100)
```

## Dwell time

### Dwell time on healthy versus unhealthy food

```{r}
#Boxplot for dwell time on each AOI in salty trials
ggplot(Salty_trials, aes(x = AOI_visited, y = DT, fill = AOI_visited)) +
  geom_boxplot() +
  labs(title = "Dwell time on healthy versus salty food",
       x = "AOI visited",
       y = "Dwell time") +
  theme_minimal() +
  ylim(0,2000) +
  theme(legend.position = "none") 

#Boxplot for dwell time on each AOI in sugary trials
xx
```

### Relative dwell time and craving

```{r message=FALSE}
# Scatter plot for Relative dwell time on unhealthy food and craving
xx

```

## Fixation count

### Number of fixations on healthy versus unhealthy food

```{r}
#Boxplot for number of fixations on each AOI in salty trials
ggplot(Salty_trials, aes(x = AOI_visited, y = n, fill = AOI_visited)) +
  geom_boxplot() +
  labs(title = "Number of fixations on healthy versus salty food",
       x = "AOI visited",
       y = "Number of fixations") +
  theme_minimal() +
  ylim(0,8) +
  theme(legend.position = "none") 

#Boxplot for number of fixations on each AOI in sugary trials
xx
```

### Relative fixation count and craving

```{r message=FALSE}
# Scatter plot for Relative fixation count on unhealthy food and craving
xx

```

# Regression analyses

## Craving predicting RT AB score

```{r}
model_RT <- lm(RT_AB_score ~ unhealthy_craving, 
                     data = All_Measures_PerPart) 

summary(model_RT)
tab_model(model_RT)
```

## Craving predicting first fixation location

```{r}
xxx
```

## Craving predicting relative dwell time

```{r}
xxx
```

## Craving predicting proportion of fixations

```{r}
xxx
```
